Community-Conditioned Language Models

### 2020-11-17 paper/paper.tex
add correlation/p-value table

We can characterise the quality the LMCC by a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given message $m$ 
from the community $c_j$ portion of the test set.  That is:
\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message.

@@

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use the entropy metric
which is has stronger information-theoretic grounds, and is a measure
of the difficulty of identifying a community, as estimated by the
model. We call this measure the (linguistic) indiscernibility of the
community $c_j$, and use \(\Ind_j\) to stand for it in formulas.

Technically, it is defined as the exponential of the
entropy of the distribution \(D_j\), which is the (discrete)
probability distribution of predicted community for a random message from
$c_j$. $D_j$ is simply the $j$th row in \(C\). In sum, we have the formula:
\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]

Indiscernibility is a robust measure: it is almost perfectly
correlated for various CCLM (transformer, lstm and various $c$) $M$.
\jp{other r's}
\begin{figure}
  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
  \caption{%
    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's r = \num{0.99} p < \num{0.0001}).
  }
  \label{fig:lmcc_ppl}
\end{figure}

However we observe that the CCLM perplexity is only weakly correlated with the
indiscernibility of the corresponding latent LMCC(\cref{fig:cclm_lmcc_ppl}).

\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean linguistic perplexity (x-axis) and community indiscernibility
    (y-axis) per community (Pearson's r = \num{0.30}, p =
    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
    other models exhibit similar behavior.  }
  \label{fig:cclm_lmcc_ppl}
\end{figure}

@@

% future work
The pattern of how much the perplexity falls across different communities
is similar across architectures, but with some differences.\jp{Either say something specific here or move the whole paragraph to future work.}
It could be interesting to further investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.

>>>>

We can characterise the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given (test set) message $m$ 
from community $c_j$. 
That is: \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).

@@

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use the entropy metric,
which is has stronger information-theoretic grounds, and is a measure
of the uncertainty of the LMCC, given messages from a particular 
community. 
Technically, the linguistic indiscernibility of community $c_j$ is defined 
as the exponential of the entropy of the $j$ row in \(C\),
which is a probability distribution 
of predicted community for a random message from $c_j$.
In sum, we have the formula: \bn{Took out $D$. is it needed elsewhere?}
\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]

Indiscernibility is a robust measure: it is almost perfectly
correlated for various CCLM (transformer, LSTM and various $c$) $M$.
In fact, we observe a correlation between the indiscernibility 
measures by every pair of models with Pearson's r $\geq 0.98$ 
and $p < 0.0001$ (see, for example, \cref{fig:lmcc_ppl}).

\begin{figure}
  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
  \caption{%
    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:lmcc_ppl}
\end{figure}

However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1. However the r is still very weakly positive for every model.}
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean community-wise linguistic perplexity (x-axis) 
    and indiscernibility (y-axis) 
    (Pearson's r = \num{0.16}; p = \num{0.2991}).  
    Shown here for the LSTM with $c = 1$.}
  \label{fig:cclm_lmcc_ppl}
\end{figure}

\begin{table}
  \centering
  \input{floats/cclm_lmcc_ppl.tex}
  \caption{
  Pearson's r correlation between community-wise linguistic perplexity and indescernibility for each model.
  }
  \label{tab:cclm_lmcc_ppl}
\end{table}

@@

The pattern of how much the perplexity falls across different communities
is similar across architectures, but with some differences.\jp{Either say something specific here or move the whole paragraph to future work.}
It could be interesting to further investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.
